Incremental Forward Reasoning for White-Box~Proof~Search

### 2025-02-19 paper.tex
Clarify addHyp

From these observations we can directly derive two procedures for adding hypotheses and matches to the rule state.
The procedures maintain the invariants of the variable indices while reporting any complete matches discovered during this process.

\begin{algorithm}
  \SetKwInOut{Precondition}{Precondition}
  \SetKwInOut{Postcondition}{Postcondition}
  \Precondition{$H$ and $M$ contain the hypotheses and matches corresponding to a context $Γ$}
  \Input{A hypothesis $h : T$ matching slot $i$ of $r$ with substitution $σ$}
  \Output{Complete matches of $r$ in context $Γ,\, h : T$ that contain $h$}
  \Postcondition{$H$ and $M$ contain the hypotheses and matches corresponding to $Γ,\, h : T$}
  \BlankLine
  \Procedure{$\mathrm{addHyp}(h : T)$}{
    \eIf{$i = 1$}{
      $\mathrm{addMatch}([h])$
    }{
      \For{$x$ \KwIn $\vars(i)$}{
        Insert $h$ into the set $H_{x}(i, σ(h))$
      }
      \KwLet $\mathit{ms} ≔ \bigcap_{x \in \vars(i)} M_{x}(i-1, σ(x))$ \\
      $\mathit{completeMs} \gets ∅$ \\
      \For{$m$ \KwIn $\mathit{ms}$}{
        $\mathit{completeMs} \gets \mathit{completeMs} ∪ \mathrm{addMatch}(m ++ [h])$
      }
      \Return{$\mathit{completeMs}$}
    }
  }
  \BlankLine % without these, the caption is not offset properly
  \BlankLine
  \caption{Adding a hypothesis to the rule state for a rule $r$}%
  \label{alg:addhyp}
\end{algorithm}

@@

% TODO pre- and postcondition?
\begin{algorithm}
  \Input{A partial match $m$ of $r$ at level $i$ with substitution $σ$}
  \Output{A set of complete matches of $r$}
  \BlankLine
  \Procedure{$\mathrm{addMatch}(h : T)$}{
    \eIf{$i$ is complete}{
      \Return{$\{m\}$}
    }{
      \For{$x$ \KwIn $\vars(i + 1)$}{
        Insert $m$ into the set $M_{x}(i, σ(x))$
      }
      \KwLet $\mathit{hs} ≔ \bigcap_{x \in \vars(i + 1)} H_{x}(i + 1, σ(x))$ \\
      $\mathit{completeMs} \gets ∅$ \\
      \For{$h$ \KwIn $\mathit{hs}$}{
        $\mathit{completeMs} \gets \mathit{completeMs} ∪ \mathrm{addMatch}(m ++ [h])$
      }
      \Return{$\mathit{completeMs}$}
    }
  }
  \BlankLine % without these, the caption is not offset properly
  \BlankLine
  \caption{Adding a match to the rule state for a rule $r$}%
  \label{alg:addmatch}
\end{algorithm}

>>>>

From these observations we can directly derive two procedures for adding hypotheses and matches to the rule state.
The procedures maintain the invariants of the variable indices while reporting any complete matches discovered during this process.
Specifically, if a rule state for a rule $r$ represents the context $Γ$---i.e., it contains the hypotheses of $Γ$ and all partial matches of $r$ that can be derived from them---then the rule state obtained from $\addHyp(h : T)$ represents $Γ,\, h : T$.

\begin{algorithm}
  \Input{A hypothesis $h : T$}
  \Output{All complete matches of $r$ in context $Γ,\, h : T$ that contain $h$}
  \BlankLine
  \Procedure{$\mathrm{addHyp}(h : T)$}{
    \For{all slots $i$ of $r$ such that $h$ matches $i$}{
      \eIf{$i = 1$}{
        $\mathrm{addMatch}([h])$
      }{
        \KwLet $σ ≔ \sub_{i}(h)$ \\
        \For{$x$ \KwIn $\vars(i)$}{
          Insert $h$ into the set $H_{x}(i, σ(h))$
        }
        \KwLet $\mathit{ms} ≔ \bigcap_{x \in \vars(i)} M_{x}(i-1, σ(x))$ \\
        $\mathit{completeMs} \gets ∅$ \\
        \For{$m$ \KwIn $\mathit{ms}$}{
          $\mathit{completeMs} \gets \mathit{completeMs} ∪ \mathrm{addMatch}(m \mathbin{++} [h])$
        }
        \Return{$\mathit{completeMs}$}
      }
    }
  }
  \BlankLine % without these, the caption is not offset properly
  \BlankLine
  \caption{Add a hypothesis to the rule state for a rule $r$ and context $Γ$}%
  \label{alg:addhyp}
\end{algorithm}

@@

% TODO pre- and postcondition?
\begin{algorithm}
  \Input{A partial match $m$ of $r$ at level $i$ with substitution $σ$}
  \Output{A set of complete matches of $r$}
  \BlankLine
  \Procedure{$\mathrm{addMatch}(h : T)$}{
    \eIf{$i$ is complete}{
      \Return{$\{m\}$}
    }{
      \For{$x$ \KwIn $\vars(i + 1)$}{
        Insert $m$ into the set $M_{x}(i, σ(x))$
      }
      \KwLet $\mathit{hs} ≔ \bigcap_{x \in \vars(i + 1)} H_{x}(i + 1, σ(x))$ \\
      $\mathit{completeMs} \gets ∅$ \\
      \For{$h$ \KwIn $\mathit{hs}$}{
        $\mathit{completeMs} \gets \mathit{completeMs} ∪ \mathrm{addMatch}(m \mathbin{++} [h])$
      }
      \Return{$\mathit{completeMs}$}
    }
  }
  \BlankLine % without these, the caption is not offset properly
  \BlankLine
  \caption{Add a match to the rule state for a rule $r$}%
  \label{alg:addmatch}
\end{algorithm}
